Beta Diversity through Time and Space

### 2020-07-16 analyses/FL_abund_GAM.Rmd
preparing to try SEM

#################################
### Join with FL and BB abund ###
#################################
bb_abund_sp_sum <- bb_abund_sp %>% # Get annual totals for BB species so that we can filter out really rare ones
  group_by(year, bb.sp) %>%
  summarize(bb.total.abund = sum(bb.abund))

@@

>>>>

write_csv(preference, "../analyses/output/preference.csv")
#################################
### Join with FL and BB abund ###
#################################
bb_abund_sp_sum <- bb_abund_sp %>% # Get annual totals for BB species so that we can filter out really rare ones
  group_by(year, bb.sp) %>%
  summarize(bb.total.abund = sum(bb.abund))

@@

Does preference relate to BB-FL traits
```{r}
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)

fl_ktype <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
  dplyr::select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_morph <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  rename(plant.sp = Species) %>%
  group_by(plant.sp) %>%
  summarize(corolla.depth = mean(LN))

fl_tax <- read_csv("../processed_data/floral_tax.csv")

pref_traits <- preference %>%
  left_join(bb_traits) %>%
  left_join(fl_ktype) %>%
  left_join(fl_morph) %>%
  separate(plant.sp, c("plant.genus", "plant.species"), sep = " ", remove = FALSE) %>%
  left_join(fl_tax) %>%
  mutate(preference.bin = case_when(
    preference > 0 ~ 1,
    preference == 0 ~ 0
  )) %>%
  mutate(tongue.mismatch = abs(pbl.w - corolla.depth),
         bb.sp = factor(bb.sp),
         plant.sp = factor(plant.sp),
         plant.genus = factor(plant.genus),
         subgenus = factor(subgenus),
         k.type.s = factor(k.type.s),
         plant.family = factor(plant.family))

ggplot(pref_traits, aes(k.type.s, preference)) +
  geom_boxplot()

ggplot(filter(pref_traits, preference > 0), aes(k.type.s, log(preference))) +
  geom_boxplot()

ggplot(pref_traits, aes(tongue.mismatch, preference)) + 
  geom_point() +
  geom_smooth()

ggplot(pref_traits, aes(abs(pbl.w - corolla.depth), preference)) + 
  geom_point() +
  geom_smooth(method = "lm")

### To deal with zero inflation, binarize and do binomial regression
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}

ggplot(pref_traits, aes(tongue.mismatch, preference.bin)) + 
  geom_point() +
  binomial_smooth()

### Then, to focus only on non-zero preference...
ggplot(filter(pref_traits, preference > 0), aes(tongue.mismatch, log(preference))) + 
  geom_point() +
  geom_smooth(method = "lm")

### Model
library(DHARMa)
lmer_nonzero <- lmer(log(preference) ~ tongue.mismatch + 
                       k.type.s +
                       (1 | plant.sp) +
                       (1 | bb.sp), 
                     data = filter(pref_traits, preference > 0 & !is.na(corolla.depth)))

summary(lmer_nonzero)
plot(lmer_nonzero)
simulationOutput <- simulateResiduals(fittedModel = lmer_nonzero, plot = T)
testResiduals(simulationOutput)
plot(simulationOutput)

lmer_withzeros <- glmer(preference ~ tongue.mismatch + 
                          k.type.s +
                          (1 | plant.sp) +
                          (1 | bb.sp), 
                       family = "poisson",
                       data = filter(pref_traits, !is.na(corolla.depth)))

summary(lmer_withzeros)
plot(lmer_withzeros)
simulationOutput_withzeros <- simulateResiduals(fittedModel = lmer_withzeros, plot = T)
testResiduals(simulationOutput_withzeros)
plot(simulationOutput_withzeros)

```

### 2020-07-16 analyses/beta_diversity.Rmd
preparing to try SEM

beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "date1"), sep = "_") %>%
  separate(j, into = c("site2", "date2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, date1, site2, date2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE)

beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE)
```

@@

### Let's look at floral survey diversity patterns
```{r}
survey_transects <- survey %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

survey_bioData <- survey %>%
  filter(flower.cover > 0) %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()
%>%
  left_join(survey_transects)

survey_sitepair <- formatsitepair(survey_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_survey <- gdm(survey_sitepair, geo = TRUE)
gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey)
plot(gdm_survey)
plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

@@

>>>>

beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)
beta_site.date_gdm.prep <- beta_site.date %>%
  filter(metric == "WN")

write_csv(beta_site.date, "../analyses/output/beta_site.date.csv")
```

### Visualize transect-level beta-diversity
```{r}
ggplot(filter(beta_site.date.2010, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2010, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2011, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2012, yday.diff == 0 &
                metric != "S" & 
                metric.class != "OS.partition"),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

ggplot(filter(beta_site.date.2012, yday.diff <= 10 &
                metric != "S" & 
                metric.class != "OS.partition"),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition" &
                site1 == site2),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

```

@@

### Let's look at floral survey diversity patterns
```{r}
### 2012
survey_bioData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2012)

### Call gdm()
gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2012)
plot(gdm_survey.2012)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

### 2011
survey_bioData.2011 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2011") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2011 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2011") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2011)

### Call gdm()
gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2011)
plot(gdm_survey.2011)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

### 2010
survey_bioData.2010 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2010") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2010 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2010") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2010)

### Call gdm()
gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2010)
plot(gdm_survey.2010)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

@@

### At the transect level
```{r}
beta_site.date_gdm.prep

transects_2010 <- beta_site.date.2010 %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev = elev1, s2.elev = elev2,
         s1.yday = yday1, s2.yday = yday2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(transects_2010, geo = TRUE)
gdm.varImp(transects_2010, geo = TRUE)
